---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# Datamations at CHI 2021

## How to run `datamations_chi2021.Rmd`

1. Make sure that `datamations_chi2021.Rmd` is in the current working directory. 
Use `setwd()` or `setwd(dirname(file.choose()))`.
2. Finally run `rmarkdown::render("datamations_chi2021.Rmd")`.

The following packages are required:

```{r, echo=FALSE}
library(tibble)
library(purrr)
library(knitr)

pkgs <- c("tidyverse", "lubridate", "knitr", "scales", "broom", "pwr", "pacman")

pkgs %>% 
  map(packageVersion) %>% 
  map_chr(as.character) %>% 
  tibble(`Package Name` = pkgs, Version = .) %>% 
  knitr::kable(format = "markdown")
```

The script uses `pacman` to manage installation of the required packages at 
these minimum version numbers. The `pacman::p_install_version` command should 
ensure that you have _at least_ these minimum version numbers, but will allow 
for newer package versions as well. If you run into difficulty when rendering 
the Rmarkdown file, it may help to manually run the following:

```{r, eval=FALSE}
install.packages("pacman")
pacman::p_install_version(
  c("tidyverse", "lubridate", "knitr", "scales", "broom", "pwr"),
  c("1.3.0", "1.7.9.2", "1.30", "1.1.1", "0.7.3", "1.3.0"))
```
