---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# Datamations at CHI 2021

This is the Github repository for "Datamations: Animated Explanations of Data Analysis Pipelines", by  
Xiaoying Pu, Sean Kross, Jake M. Hofman, and Daniel G. Goldstein, published at CHI '21: Proceedings of the 2021 CHI Conference on Human Factors in Computing Systems.

It includes everything needed to reproduce the figures and results in the paper.

The paper is available online ([publisher link](https://doi.org/10.1145/3411764.3445063), [pdf](http://jakehofman.com/pdfs/datamations.pdf)) and supplementary material is available [here](https://osf.io/85njc/).

If you would like to use Datamations yourself, please see this Github repository for the latest code: https://github.com/microsoft/datamations

## How to run `datamations_chi2021.Rmd`

1. Make sure that `datamations_chi2021.Rmd` is in the current working directory. 
Use `setwd()` or `setwd(dirname(file.choose()))`.
2. Finally run `rmarkdown::render("datamations_chi2021.Rmd")`.

The following packages are required:

```{r, echo=FALSE}
library(tibble)
library(purrr)
library(knitr)

pkgs <- c("tidyverse", "lubridate", "knitr", "scales", "broom", "pwr", "pacman")

pkgs %>% 
  map(packageVersion) %>% 
  map_chr(as.character) %>% 
  tibble(`Package Name` = pkgs, Version = .) %>% 
  knitr::kable(format = "markdown")
```

The script uses `pacman` to manage installation of the required packages at 
these minimum version numbers. The `pacman::p_install_version` command should 
ensure that you have _at least_ these minimum version numbers, but will allow 
for newer package versions as well. If you run into difficulty when rendering 
the Rmarkdown file, it may help to manually run the following:

```{r, eval=FALSE}
install.packages("pacman")
pacman::p_install_version(
  c("tidyverse", "lubridate", "knitr", "scales", "broom", "pwr"),
  c("1.3.0", "1.7.9.2", "1.30", "1.1.1", "0.7.3", "1.3.0"))
```
